# Cleaning my data

## Overview of the Chapter
The first step was to clean all my data. The values collected in this experiment were percent cover of each species present in my 1m plot. Although the data are technically in cover classes (<1, 1-10, 10-20, 20-30, 30-40, 40-50, 50-60, 60-70, 70-80, 80-90, >99), data were written down as a single number meant to represent the lower value of the cover class (<1, 10, 20, 30, 40, 50, 60, 70, 80, 90, >99). For my data analyses, I will use the mid-point of the cover classes (5, 15, 25, 35, 45, 55, 65, 75, 85, 95). I intend to use the beta distribution for my analyses, and therefore need to nudge my upper and lower bounds into values between 0 and 1 (<1 -> 0.005, >99 -> .999). 

## Importing the data
I have two different spreadsheets, each representing one of my field sites. Both will be cleaned in the same way, but first I need to import the data and load all my packages. 

```{r data}
#Load packages
library(tidyverse)
library(lubridate) #my version of tidyverse does not seem to include lubridate

#Import data
fb <- read.csv("../../Cleaned-Data/clean_fb.csv")
ul <- read.csv("../../Cleaned-Data/clean_ul.csv")

```

## Cleaning the data
To begin, I will start by working with `fb`, which is all my data from Farmington Bay. One of the columns in my table is Date, which is the date of sampling.

```{r class}

class(fb$Date)

```

R is reading my date column as a character string! I need to fix this using lubridate so that R understands that my dates are dates.

```{r date}

fb$Date <- lubridate::mdy(fb$Date)

class(fb$Date)
```

Now I can move on to fixing my column names. When I `glimpse` my column names, I can see that not all of them make sense.

```{r glimpse}

dplyr::glimpse(fb)

```
My first column name is coming up as Block...FB. This column name does not follow the conventions we learned in class and it too complicated to type. I will change it to simply say Block. I also need to change the column Unk_Rush to be called Unk_Bulrush

```{r block}

names(fb)[1] <- "Block"
names(fb)[25] <- "Unk_Bulrush"

```

The most difficult part of cleaning my data is making sure all the cover values have been entered properly. Let's start by looking at the values in a random column.

```{r test}

unique(fb$BOMA)

```
We have a couple of problems here. First of all, the values are all being read as characters instead of numbers. This is because of my values that reads "<1". My first step will be to replace this value with a trace value that can be used in statistical analyses (typically one half of the smallest recorded value). After replacing that value, I will be able to change all these values into numeric values so they are no longer characters. This will also help to change my "" values to NA values. Next, my larger values were entered into the spreadsheet as percentages (10 = 10%, 20 = 20%). To perform my statistical analyses, these will need to be changes into proportions (10 -> .1, 20 -> .2). Finally, I want all my values to be on the mid-point of their cover class, not the bottom. Therefore, everything needs to increase by 0.05 (.1 -> .15, .2 -> .25).

```{r mutate}

fb <- fb %>% 
  mutate(Total.Cover = case_when(
  Total.Cover == "<1" ~ 0.005, #change all the <1 to the trace value of 0.005
  Total.Cover == "1" ~ 0.05, #change the 1s to 0.05 (because too small to work below)
  #all other numbers can stay as they are but numeric
  #but add 5 so they are at the mid-point of the cover class
  #and then divide them by 100 so they are proportions
  TRUE ~ (as.numeric(Total.Cover) + 5)/100
))

```

Now I just need to repeat this for all the columns between Total.Cover and Tamarisk. I originally tried to do this through a loop so I wouldn't need to copy and paste for each column, but unfortunately I could not get the loop to run in time. That can be a future project for me!

```{r all}

fb <- fb %>% 
  mutate(PHAU = case_when(
          PHAU == "<1" ~ 0.005, 
          PHAU == "1" ~ 0.05, 
          TRUE ~ (as.numeric(PHAU) + 5)/100 ),
         Cheno = case_when(
          Cheno == "<1" ~ 0.005, 
          Cheno == "1" ~ 0.05, 
          TRUE ~ (as.numeric(Cheno) + 5)/100),
        Typha = case_when(
          Typha == "<1" ~ 0.005, 
          Typha == "1" ~ 0.05, 
          TRUE ~ (as.numeric(Typha) + 5)/100),
        BOMA = case_when(
          BOMA == "<1" ~ 0.005, 
          BOMA == "1" ~ 0.05, 
          TRUE ~ (as.numeric(BOMA) + 5)/100),
        DISP = case_when(
          DISP == "<1" ~ 0.005, 
          DISP == "1" ~ 0.05, 
          TRUE ~ (as.numeric(DISP) + 5)/100),
        EUMA = case_when(
          EUMA == "<1" ~ 0.005, 
          EUMA == "1" ~ 0.05, 
          TRUE ~ (as.numeric(EUMA) + 5)/100),
        SYCI = case_when(
          SYCI == "<1" ~ 0.005, 
          SYCI == "1" ~ 0.05, 
          TRUE ~ (as.numeric(SYCI) + 5)/100),
        LEFA = case_when(
          LEFA == "<1" ~ 0.005, 
          LEFA == "1" ~ 0.05, 
          TRUE ~ (as.numeric(LEFA) + 5)/100),
        SCAC = case_when(
          SCAC == "<1" ~ 0.005, 
          SCAC == "1" ~ 0.05, 
          TRUE ~ (as.numeric(SCAC) + 5)/100),
        BICE = case_when(
          BICE == "<1" ~ 0.005, 
          BICE == "1" ~ 0.05, 
          TRUE ~ (as.numeric(BICE) + 5)/100),
        BIFR = case_when(
          BIFR == "<1" ~ 0.005, 
          BIFR == "1" ~ 0.05, 
          TRUE ~ (as.numeric(BIFR) + 5)/100),
        EUOC = case_when(
          EUOC == "<1" ~ 0.005, 
          EUOC == "1" ~ 0.05, 
          TRUE ~ (as.numeric(EUOC) + 5)/100),
        MUAS = case_when(
          MUAS == "<1" ~ 0.005, 
          MUAS == "1" ~ 0.05, 
          TRUE ~ (as.numeric(MUAS) + 5)/100),
        SCAM = case_when(
          SCAM == "<1" ~ 0.005, 
          SCAM == "1" ~ 0.05, 
          TRUE ~ (as.numeric(SCAM) + 5)/100),
        RUMA = case_when(
          RUMA == "<1" ~ 0.005, 
          RUMA == "1" ~ 0.05, 
          TRUE ~ (as.numeric(RUMA) + 5)/100),
        RUST = case_when(
          RUST == "<1" ~ 0.005, 
          RUST == "1" ~ 0.05, 
          TRUE ~ (as.numeric(RUST) + 5)/100),
        Unk_Forb = case_when(
          Unk_Forb == "<1" ~ 0.005, 
          Unk_Forb == "1" ~ 0.05, 
          TRUE ~ (as.numeric(Unk_Forb) + 5)/100),
        Unk_Grass = case_when(
          Unk_Grass == "<1" ~ 0.005, 
          Unk_Grass == "1" ~ 0.05, 
          TRUE ~ (as.numeric(Unk_Grass) + 5)/100),
        Unk_Bulrush = case_when(
          Unk_Bulrush == "<1" ~ 0.005, 
          Unk_Bulrush == "1" ~ 0.05, 
          TRUE ~ (as.numeric(Unk_Bulrush) + 5)/100),
        SARU = case_when(
          SARU == "<1" ~ 0.005, 
          SARU == "1" ~ 0.05, 
          TRUE ~ (as.numeric(SARU) + 5)/100),
        Tamarisk = case_when(
          Tamarisk == "<1" ~ 0.005, 
          Tamarisk == "1" ~ 0.05, 
          TRUE ~ (as.numeric(Tamarisk) + 5)/100)
  )
```

Now we can go in and do the same to the `ul` spreadsheet. 

Once we are done, the next step is visualizing our data!